### 2020-05-03 Case Studies/03_spam_identification.rmd
spam identification using statistics
spam identification using statistics

```{r}
head(list.files(path = file.path(data.dir, "easy_ham")))
head(list.files(path = file.path(data.dir, "spam_2")))

@@

```{r}
getBoundary <- function(header) {

   boundaryIdx <- grep("boundary=", header)
   boundary = gsub('"', "", header[boundaryIdx])
   gsub(".*boundary= *([^;]*);?.*", "\\1", boundary)
}

@@

```{r}
sampleSplit[[6]]

@@

msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
tail(msg)
```

@@

dropAttach <- function(email) {

}

@@

```{r}
stopWords <- stopwords(kind = "en")
```

@@

words <- words[ ! (words %in% stopWords) ]

@@

findMsgWords <- function(msg, stopwords = stopWords) {

@@

   words <- words[ ! (words %in% stopwords) ]

   return(words)   
}

findMsgWords(msg)
```

@@

```{r}
processAllWords <- function(dirName, stopWords) {
   # read all files in the directory
   fileNames <- list.files(dirName, full.names = T)

@@

@@

msgWordList <- lapply(directories, processAllWords)

@@

>>>>

head(list.files(path = file.path(data.dir, "easy_ham")))
head(list.files(path = file.path(data.dir, "spam_2")))

@@

getBoundary <- function(header) {
   boundaryIdx <- grep("boundary=", header)
   boundary = gsub('"', "", header[boundaryIdx])
   gsub(".*boundary= *([^;]*);?.*", "\\1", boundary)
}

@@

@@

msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
tail(msg)

@@

@@

```{r}
exclude_word_list <- stopwords(kind = "en")
```

@@

words <- words[ ! (words %in% exclude_word_list) ]

@@

findMsgWords <- function(msg, exclude) {

@@

   keep <- sapply(words, function(word) return(!(word %in% exclude)))

   return(words[ keep ])   
}

@@

dropAttach <- function(body, boundary) {

   if(is.null(body)) {
      return("")
   }

   bString <- paste("--", boundary, sep = "")
   bStringLocs <- which(bString == body)

   eString <- paste("--", boundary, "--", sep = "")
   eStringLoc <- which(eString == body)

   if(length(bStringLocs) == 2) {
      msg <- body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)]
   }

   if(length(eStringLoc) > 0) {
      msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
   }

   return(msg)
}

processAllWords <- function(dirName, stopWords) {
   # read all files in the directory
   fileNames <- list.files(dirName, full.names = T)

@@

   # find boundary string for messages with attachments
   boundaries <- sapply(headerList[hasAttach], getBoundary)

@@

msgWordList <- lapply(directories, processAllWords, stopWords = exclude_word_list)

numMsgs <- sapply(msgWordList, length)
numMsgs

isSpam <- rep(c(FALSE, FALSE, FALSE, TRUE, TRUE), numMsgs)

msgWordsList <- unlist(msgWordList, recursive = F)

@@

## Naive Bayes Classifier Implementation

### Train / Test Split

```{r}
numEmail <- length(isSpam)

numSpam <- sum(isSpam)
numHam <- numEmail - numSpam

set.seed(418910)

testSpamIdx <- sample(numSpam, size = floor(numSpam/3))
testHamIdx <- sample(numHam, size = floor(numHam/3))

testMsgWords <- c((msgWordsList[isSpam])[testSpamIdx],
                  (msgWordsList[!isSpam])[testHamIdx])

trainMsgWords <- c((msgWordsList[isSpam])[ - testSpamIdx ],
                   (msgWordsList[!isSpam])[ - testHamIdx])

testIsSpam <- rep(c(T, F), 
                  c(length(testSpamIdx), length(testHamIdx)))

trainIsSpam <- rep(c(T, F),
                   c(numSpam - length(testSpamIdx),
                     numHam - length(testHamIdx)))

### Probability Estimates from Training Sample

```{r}
bow <- unique(unlist(trainMsgWords))

length(bow)
```

```{r}
spamWordCounts <- rep(0, length(bow))

names(spamWordCounts) = bow

tmp <- lapply(trainMsgWords[trainIsSpam], unique)
tt <- table( unlist(tmp) )
spamWordCounts[ names(tt) ] = tt

spamWordsProbs <- (spamWordCounts + 0.5) / (sum(trainIsSpam) + 0.5)

spamWordsProbs[1:20]

hamWordCounts <- rep(0, length(bow))

names(hamWordCounts) = bow

tmp <- lapply(trainMsgWords[ - trainIsSpam], unique)
tt <- table( unlist(tmp) )
hamWordCounts[ names(tt) ] = tt

hamWordsProbs <- (hamWordCounts + 0.5) / (sum(!trainIsSpam) + 0.5)

log(spamWordsProbs) - log(hamWordsProbs)

```{r}
wordsList <- trainMsgWords
spam <- trainIsSpam

make_words_valid_columns <- function( words, all_words ) {

   word_counts <- rep(0, length(all_words))
   names(word_counts) <- all_words

   tmp <- lapply(words, unique)
   tt <- table( unlist(tmp) )
   word_counts[ names(tt) ] = tt

   return(word_counts)
}

computeFreqs <- function(wordsList, spam, bow = unique(unlist(wordsList))) {

   all_words <- unique(bow)

   # create a matrix for spam, ham, and log odds
   wordTable <- matrix(0.5, nrow = 2, ncol = length(bow))
   colnames(wordTable) <- all_words
   rownames(wordTable) <- c( "presentLogOdds",
                          "absentLogOdds")

   # for each spam message, add 1 to the counts for words in messsage

   spam_all <- wordsList[spam]
   spam_words <- make_words_valid_columns( spam_all, all_words )

   wordTable <- rbind(wordTable, spam_words + 0.5)
   rownames(wordTable)[3] <- "spam"

   # Similarly for ham messages

   ham_all <- wordsList[ !spam ]

   ham_words <- make_words_valid_columns( ham_all, all_words )

   wordTable <- rbind(wordTable, ham_words + 0.5)
   rownames(wordTable)[4] <- "ham"

   head(wordTable[, 1:20])

   # find the total number of spam and ham
   numSpam <- sum(spam)
   numHam <- length(spam) - numSpam

   # prob (word|spam) and prob(words|ham)
   wordTable["spam", ] <- wordTable["spam", ] / (numSpam + 0.5)
   wordTable["ham", ] <- wordTable["ham", ] / (numHam + 0.5)

   head(wordTable[, 1:20])

   # log odds
   wordTable["presentLogOdds", ] = 
      log(wordTable["spam", ]) - log(wordTable["ham", ])

   wordTable["absentLogOdds", ] =
      log((1 - wordTable["spam", ])) - log((1 - wordTable["ham", ]))

   invisible(wordTable)
}

```{r}
trainTable <- computeFreqs(trainMsgWords, trainIsSpam)

# peek the prob table
head(trainTable[, 1:10])
```

### Classifying New Messages

```{r}
newMsg <- testMsgWords[[1]]

# only look at words we have classified
newMsg <- newMsg[ !is.na(match(newMsg, colnames(trainTable)))]

present <- colnames(trainTable) %in% newMsg

sum( trainTable["presentLogOdds", present]) +
   sum( trainTable["absentLogOdds", !present])

newMsg <- testMsgWords[[ which(!testIsSpam)[ 1 ] ]]
newMsg <- newMsg[ !is.na(match(newMsg, colnames(trainTable)))]
present <- (colnames(trainTable) %in% newMsg)

sum(trainTable["presentLogOdds", present]) +
   sum(trainTable["absentLogOdds", !present])
